Combinatorial Proofs and Decomposition Theorem for First-order Logic

### 2021-01-06 main.tex
line numbers and page numbers
Fix the macros and reorganize the structure

@@

\newcommand{\ax}{\mathsf{ax}}
\newcommand{\cut}{\mathsf{cut}}
\newcommand{\conj}{\mathsf{\wedge}}
\newcommand{\disj}{\mathsf{\vee}}
\newcommand{\univ}{\mathsf{\forall}}
\newcommand{\exist}{\mathsf{\exists}}
\newcommand{\mix}{\mathsf{mix}}
% ** SPECIAL **

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{On First-order Proofs without syntax}

@@

\section{First-order Combinatorial Proofs}
\subsection{First-order Logic and Graphs}

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$\delta$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $\delta t_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\vee$ and $\wedge$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is \bfit{rectified} if all bound variables are distinct from one another and from all free variables. Without loss of generality, we assume that all formulas are rectifed throughtout the paper as every formula has a logically equivalent rectified form.

@@

For a formula $A$, we can define its associated fograph $\Gr(A)$ inductively by: $\Gr(\alpha) = \cdot \alpha$ (a single-vertex graph whose vertex is labelled by $\alpha$); $\Gr(A \vee B) = \Gr(A) + \Gr(B)$ and $\Gr(A \wedge B) = \Gr(A) \times \Gr(B)$.
The \bfit{binding graph} of a forgraph $G$ is the directed graph $(V_G, \{(b,
l) | b \text{ binds } l\})$.

@@

In $\FOMLL$, terms, atoms, formulas are defined as in first-order logic. For
simplicity, we choose to use $\vee$ and $\wedge$ instead of $\parr$ and
$\otimes$ which are generally used in the presentation of linear logic.
A formula $A$ is identified with its \bfit{formula tree} $\FT(A)$, a directed
tree with leaves labelled by atoms, internal nodes labelled by connectives and
quantifiers, and edges directed from leaves to the root. A \bfit{sequent} $\Gamma$ is simply a disjoint union of formulas. We write
comma for disjoint union.

@@

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\conj$]{\vdash \Gamma, \Delta, A \wedge B}
\end{prooftree}
\hspace{1.5cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\disj$]{\vdash \Gamma, A \vee B}
\end{prooftree}}

@@

Let $\lambda$ is a unifiable linking on $\Gamma = A_1, \ldots, A_n$.
Let $A = A_1 \vee \ldots \vee A_n$. The \bfit{unification structure}
$\US(\lambda)$
associated to $\lambda$ is the formula tree $\FT(A)$ together with an undirected
edge between leaves $l$ and $l'$ for every link $\{l, l'\}$ in $\lambda$ and a
directed edge from $\exists x$ to $\forall y$ for every dependency $\exists x
\rightarrow \forall y$.

A \bfit{switching graph} of a unification structure $\US(\lambda)$ is any
derivative of $\US(\lambda)$ obtained by keeping only one edge into each $\vee$
and $\forall$ and undirecting remaining edges. A linking is \bfit{correct} if it
is unifiable and all of the switching graphs of its associated unification
structure are acyclic.

@@

\subsection{Deep Inference}
In this paper we also use some \bfit{deep inference} \cite{BT 2001} rules that are in the
following form:
\begin{center}
\begin{prooftree}
  \hypo{\vdash S\{A\}}
  \infer1[]{\vdash S\{B\}}
\end{prooftree}
\end{center}
where $S\{ \ \}$ stands for a \bfit{context}, which corresponds to a sequent
with a hole taking the place of an atom, and $S\{A\}$ represents
the sequent or formula obtained by replacing the hole in $S\{ \ \}$ with the formula $A$. Formally,

  \[C ::= \Box \ | \ A \vee C \ | \ C \wedge A \ | \ \exists x C \ | \ \forall x C \].
  \[S ::= C \ | \ A, S \ | \ S, A \] where $A$ is a formula.
The above rule can be thus seen as the rewriting rule $A \rightarrow B$.

@@

\subsection{$\LK$}
We start with Gentzen's sequent calculus system $\LK$ \cite{Gentzen 1935}.

\mbox{
\begin{prooftree}
  \infer0[$ax$]{\vdash A, \neg A}
\end{prooftree}
\hspace{2cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, \neg A}
  \infer2[$cut$]{\vdash \Gamma, \Delta}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\wedge$]{\vdash \Gamma, \Delta, A \wedge B}
\end{prooftree}
\hspace{1.5cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\vee$]{\vdash \Gamma, A \vee B}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$ctr$]{\vdash \Gamma, A}
\end{prooftree}
	\hspace{2cm}

\begin{prooftree}
  \hypo{\vdash \Gamma}
  \infer1[$wk$]{\vdash \Gamma, A}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \infer1[$\forall$]{\vdash \Gamma, \forall x A}
\end{prooftree}
($x \notin fv(\Gamma)$)
\hspace{2cm}

Note that the $\mix$ rule is derivable via weakening in $\LK$.

Consider the following deep inference rules \cite{BT 2001}:
\begin{center}
\mbox{
\begin{prooftree}
  \hypo{\vdash S\{A \vee A\}}
  \infer1[$\cD$]{\vdash S\{A\}}
\end{prooftree}
\hspace{2cm}

@@

Note that the $ctr$ (resp. $wk$) rule in $\LK$ is derivable in $\{\cD, \vee\}$
(resp. $\{\wD, \fff\}$) and that $\cD$ and $\wD$ rules
permute downwards with the non-structural rules of $\LK$.

@@

\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$ctr$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$\vee$]{\vdash \Gamma, A \vee A}
  \infer1[$\cD$]{\vdash \Gamma, A}
\end{prooftree}}
\vspace{0.2cm}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma}
  \infer1[$wk$]{\vdash \Gamma, A}
\end{prooftree}

@@

\mbox{
\begin{prooftree}
\hypo{\Gamma,A \vee A}
\infer1[$\cD$]{\Gamma,A} 
\hypo{\Delta,B}
\infer2[$\wedge$]{\Gamma, \Delta, A \wedge B} 
\end{prooftree}

@@

\begin{prooftree}
\hypo{\Gamma,A \vee A}
\hypo{\Delta,B}
\infer2[$\wedge$]{\Gamma, \Delta, (A \vee A) \wedge B} 
\infer1[$\cD$]{\Gamma, \Delta, A \wedge B} 
\end{prooftree}}

@@

($\Leftarrow$)
We regard the proof in $\FOMLLm$ as a proof in $\LKm$.
Then we put the derivation consisting of only $\cD$ and $\wD$ under the proof in $\LKm$. Now we try to permute all the instances $\cD$ and $\wD$ upwards with the rules of LK and $mix$. 
For the $\cD$ part, the only non-trivial case is the permutation with the $\vee$ rule where the formula generated is $A \vee A$.

\begin{center}
\mbox{
\begin{prooftree}
\hypo{\vdash \Gamma, A, A}
\infer1[$\vee$]{\vdash \Gamma, A \vee A}
\infer1[$\cD$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{prooftree}
\hypo{\vdash \Gamma}
\infer1[$wk$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{definition} A \textit{switching path} of a unification structure $U(\lambda)$ is a path in a switching graph of $U(\lambda)$.
\end{definition}

\begin{definition} A \textit{switching path} of a  formula tree $\FT(A)$ is a
	path in $\FT(A)$ that does not go through both incoming edges of a $\vee$.
\end{definition}

@@

\begin{proposition}
\label{prop2}
Let $u$ and $v$ be two distinct vertices of a fograph $\Gr(A)$, then we have the equivalence
between:
\begin{itemize}
  \item $u$ and $v$ are adjacent in $\Gr(A)$
  \item $u$ and $v$ are connected by a switching path of $F(A)$, and if one of them is a universal quantifier, then the other is
	not a descendant of the former.
  \end{itemize}
\begin{proof}
By induction on $A$.
\begin{itemize}
  \item If $A$ is an atom, trivial.
  \item If $A = A_1 \wedge A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are
	      adjacent in $\Gr(A)$ by definition. By Proposition \ref{prop1},
		    the one in $A_1$ (resp. $A_2$) is connected to the
		    vertex representing $A_1$ (resp. $A_2$) by a switching
		    path. Together with the two edges incident to $A_1 \wedge
		    A_2$, we obtain a switching path connecting $u$ and $v$.
    \end{itemize}
  \item If $A = A_1 \vee A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are not
	      adjacent in $\Gr(A)$ by definition. It is clear that they
		    are not connected by a switching path.
    \end{itemize}
  \item If $A = \exists x \ A'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A'$: trivial by the induction
	      hypothesis.
      \item one of them is $\exists x$ and the other is in $A'$: trivial
	      by Proposition \ref{prop1}

@@

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = \Gr(A)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $A$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\vee$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\vee$: we can choose an
	    arbitrary switching for this $\vee$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching
		  path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes
		  through the left (resp. right) edge of $\vee$.

    \begin{center}	
    \begin{tikzpicture}
    \tikzstyle{every node}=[circle, minimum size=12pt, inner sep=2pt]
    \draw (0, 0) node (par) {$\vee$};
    \draw (1, 1.7) node (dots1) {$\vdots$};
    \draw (-1, 1.7) node (dots2) {$\vdots$};
    \draw (1, 3.4) node (vr) {$v_R$};
    \draw (-1, 3.4) node (vl) {$v_L$};
    \draw (1.5, -1) node (dots3) {$\vdots$};
    \draw (-1.5, -1) node (dots4) {$\vdots$};
    \draw (2.9, -1) node (ul) {$u_L$};
    \draw (-2.9, -1) node (ur) {$u_R$};
    \draw (par) -- (dots1) [color=blue];
    \draw (par) -- (dots2) [color=red];
    \draw (dots1) -- (vr) [color=blue]; 
    \draw (dots2) -- (vl) [color=red];
    \draw (dots4) -- (ur) [color=blue]; 
    \draw (dots3) -- (ul) [color=red];
    \draw (dots4) -- (par) [color=blue]; 
    \draw (dots3) -- (par) [color=red];
    \end{tikzpicture}

    {\bf Figure 7.} A schema showing that the two branches of the same $\vee$ cannot be used in the cycle at the same time. 
    \end{center}

@@

    It is clear that $u_L$ (resp. $u_R$) is not in the branches of the
	 $\vee$. Otherwise, there will be no switching path from $u_L$ to $v_L$

    By Proposition \ref{prop2}, we know that $u_L$ and $u_R$ are not
	universal quantifiers which are ancestors the $\vee$
	and that there exist one switching path from $u_L$ to $v_L$ and one
	from $u_R$ to $v_R$. In particular, there exist one switching path from
	$u_L$ to the $\vee$ and one from the $\vee$ to $v_R$, and by
	concatenating the two, we obtain a switching path from $u_L$ to $v_R$. 
	By Proposition \ref{prop2}, $u_L$ and $v_R$ are thus adjacent in $(W,
	E_G)$, which is impossible since $(W, E_W)$ is a matching.
  \end{itemize}
  Notice that the switching paths here are in the underlying formula tree. We
  have to verify that they are compatible with the choices of switching made for
  universal quantifiers. That is, if $uv \in E_W$, then for all the
  universal quantifiers $\forall x$ on the switching path (in the formula
	tree), we have chosen in the first part of the proof to keep the only
	edge from the child of $\forall x$ to itself. In fact, if there exists
	a universal quantifier $w \in W$ on the switching path $u \rightarrow
	v$, then one of $u$ and $v$ is not a descendant of $w$. Moreover, if
	$u$ (resp. $v$) is a universal quantifier, then $w$ is not in its scope.
	By Proposition \ref{prop2}, $\{wu, wv\} \cap E_W \neq \emptyset$, which
	is impossible since $(W, E_W)$ is a matching.
  We have thus constructed a switching graph containing this cycle.
\end{proof}
\end{proposition}

@@

\begin{definition}
Let $\theta$ be a unification structure on an $\FOMLL$ sequent $\Gamma$.
We define the \textit{frame}  of $\theta$ by exhaustively applying the following subformula rewriting steps, to obtain a proof structure $\theta_m$ on an $\MLL$ sequent $\Gamma_m$:

\begin{enumerate}
  \item {\bf Encode dependencies as fresh links.} For each dependency $\exists x
	  \rightarrow \forall y$, with corresponding subformulas $\exists x A$
		and $\forall y B$, we add a fresh link as follows. Let $P$ be a
		fresh (nullary) predicate symbol. Replace $\exists x A$ with $P
		\wedge \exists x A$ and $\forall y B$ with $\overline{P} \vee \forall y B$, and add an axiom link between $P$ and $\overline{P}$.
  \item {\bf Erase quantifiers.} After step 1, erase all the quantifiers. (We no longer need their leaps since they are encoded as links in step 1.
  \item {\bf Simplify atoms.} After step 2, replace every predicate $Pt_1 \cdots t_n$ with a nullary predicate symbol $P$.
\end{enumerate}
\end{definition}

>>>>

% ** RULES **

\newcommand{\ax}{\mathsf{ax}}
\newcommand{\cut}{\mathsf{cut}}
\newcommand{\conj}{\mathsf{\wedge}}
\newcommand{\disj}{\mathsf{\vee}}
\newcommand{\univ}{\mathsf{\forall}}
\newcommand{\exist}{\mathsf{\exists}}
\newcommand{\mix}{\mathsf{mix}}
\newcommand{\ctr}{\mathsf{ctr}}
\newcommand{\wk}{\mathsf{wk}}

@@

\newcommand{\cor}{\vee}
\newcommand{\cand}{\wedge}

% ** SPECIAL **

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Combinatorial Proofs and Decomposition Theorem for First-order Logic}

@@

\section{First-order Combinatorial Proofs}
\subsection{First-order Logic}

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$\delta$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $\delta t_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\cor$ and $\cand$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is
\bfit{rectified} if all bound variables are distinct from one another and from
all free variables. Without loss of generality, we assume that all formulas are
rectifed throughtout the paper as every formula has a logically equivalent
rectified form. A formula is \bfit{semi-rectified} if it can be rectified in a
unique way via variable renaming.

We consider Gentzen's sequent calculus system $\LK$ \cite{Gentzen 1935}.

A \bfit{sequent} is a multiset of formulas and sequents are proved using the
following inference rules.
\begin{center}

\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, \neg A}
  \infer2[$\cut$]{\vdash \Gamma, \Delta}
\end{prooftree}}
\vspace{0.2cm}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\cand$]{\vdash \Gamma, \Delta, A \cand B}
\end{prooftree}
\hspace{1.5cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\cor$]{\vdash \Gamma, A \cor B}
\end{prooftree}}
\vspace{0.2cm}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$\ctr$]{\vdash \Gamma, A}
\end{prooftree}
	\hspace{2cm}

\begin{prooftree}
  \hypo{\vdash \Gamma}
  \infer1[$\wk$]{\vdash \Gamma, A}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \infer1[$\forall$]{\vdash \Gamma, \forall x A}
\end{prooftree}
($x \notin fv(\Gamma)$)
\hspace{1.2cm}

In this paper, we also use some \bfit{deep inference} \cite{BT 2001} rules that are in the
following form:
\begin{center}
\begin{prooftree}
  \hypo{\vdash S\{A\}}
  \infer1[]{\vdash S\{B\}}
\end{prooftree}
\end{center}
where $S\{ \ \}$ stands for a \bfit{context}, which corresponds to a sequent
with a hole taking the place of an atom, and $S\{A\}$ represents
the sequent or formula obtained by replacing the hole in $S\{ \ \}$ with the formula $A$. Formally,

  \[C ::= \Box \ | \ A \cor C \ | \ C \cand A \ | \ \exists x C \ | \ \forall x C \].
  \[S ::= C \ | \ A, S \ | \ S, A \] where $A$ is a formula.
The above rule can be thus seen as the rewriting rule $A \rightarrow B$.

\subsection{Graphs}

@@

For a formula $A$, we can define its associated fograph $\Gr(A)$ inductively by: $\Gr(\alpha) = \cdot \alpha$ (a single-vertex graph whose vertex is labelled by $\alpha$); $\Gr(A \cor B) = \Gr(A) + \Gr(B)$ and $\Gr(A \cand B) = \Gr(A) \times \Gr(B)$.
The \bfit{binding graph} of a forgraph $G$ is the directed graph $(V_G, \{(b,
l) | b \text{ binds } l\})$.

@@

In $\FOMLL$, terms, atoms, formulas are defined as in first-order logic. For
simplicity, we choose to use $\cor$ and $\cand$ instead of $\parr$ and
$\otimes$ which are generally used in the presentation of linear logic.
A formula $A$ is identified with its \bfit{formula tree} $\FT(A)$, a directed
tree with leaves labelled by atoms, internal nodes labelled by connectives and
quantifiers, and edges directed from leaves to the root. A \bfit{sequent} $\Gamma$ is simply a disjoint union of formulas. We write
comma for disjoint union.

@@

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\conj$]{\vdash \Gamma, \Delta, A \cand B}
\end{prooftree}
\hspace{1.5cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\disj$]{\vdash \Gamma, A \cor B}
\end{prooftree}}

@@

Let $\lambda$ is a unifiable linking on $\Gamma = A_1, \ldots, A_n$.
Let $A = A_1 \cor \ldots \cor A_n$. The \bfit{unification structure}
$\US(\lambda)$
associated to $\lambda$ is the formula tree $\FT(A)$ together with an undirected
edge between leaves $l$ and $l'$ for every link $\{l, l'\}$ in $\lambda$ and a
directed edge from $\exists x$ to $\forall y$ for every dependency $\exists x
\rightarrow \forall y$.

A \bfit{switching graph} of a unification structure $\US(\lambda)$ is any
derivative of $\US(\lambda)$ obtained by keeping only one edge into each $\cor$
and $\forall$ and undirecting remaining edges. A linking is \bfit{correct} if it
is unifiable and all of the switching graphs of its associated unification
structure are acyclic.

@@

@@

\subsection{Decomposition Theorem}
Consider the following deep inference rules \cite{BT 2001}:
\begin{center}
\mbox{
\begin{prooftree}
  \hypo{\vdash S\{A \cor A\}}
  \infer1[$\cD$]{\vdash S\{A\}}
\end{prooftree}
\hspace{2cm}

@@

Note that the $\ctr$ (resp. $\wk$) rule in $\LK$ is derivable in $\{\cD, \cor\}$
(resp. $\{\wD, \fff\}$) and that $\cD$ and $\wD$ rules
permute downwards with the non-structural rules of $\LK$.

@@

\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$\ctr$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$\cor$]{\vdash \Gamma, A \cor A}
  \infer1[$\cD$]{\vdash \Gamma, A}
\end{prooftree}}
\vspace{0.2cm}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma}
  \infer1[$\wk$]{\vdash \Gamma, A}
\end{prooftree}

@@

\mbox{
\begin{prooftree}
\hypo{\Gamma,A \cor A}
\infer1[$\cD$]{\Gamma,A} 
\hypo{\Delta,B}
\infer2[$\cand$]{\Gamma, \Delta, A \cand B} 
\end{prooftree}

@@

\begin{prooftree}
\hypo{\Gamma,A \cor A}
\hypo{\Delta,B}
\infer2[$\cand$]{\Gamma, \Delta, (A \cor A) \cand B} 
\infer1[$\cD$]{\Gamma, \Delta, A \cand B} 
\end{prooftree}}

@@

($\Leftarrow$)
We regard the proof in $\FOMLLm$ as a proof in $\LKm$.
Then we put the derivation consisting of only $\cD$ and $\wD$ under the proof in $\LKm$. Now we try to permute all the instances $\cD$ and $\wD$ upwards with the rules of LK and $mix$. 
For the $\cD$ part, the only non-trivial case is the permutation with the $\cor$ rule where the formula generated is $A \cor A$.

\begin{center}
\mbox{
\begin{prooftree}
\hypo{\vdash \Gamma, A, A}
\infer1[$\cor$]{\vdash \Gamma, A \cor A}
\infer1[$\cD$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{prooftree}
\hypo{\vdash \Gamma}
\infer1[$\wk$]{\vdash \Gamma, A}
\end{prooftree}

@@

\begin{definition} A \bfit{switching path} of a unification structure $U(\lambda)$ is a path in a switching graph of $U(\lambda)$.
\end{definition}

\begin{definition} A \bfit{switching path} of a  formula tree $\FT(A)$ is a
	path in $\FT(A)$ that does not go through both incoming edges of a $\cor$.
\end{definition}

@@

\begin{proposition}
\label{prop2}
Let $u$ and $v$ be two distinct vertices of a fograph $\Gr(A)$, then we have the equivalence
between:
\begin{itemize}
  \item $u$ and $v$ are adjacent in $\Gr(A)$
  \item $u$ and $v$ are connected by a switching path of $\FT(A)$, and if one of them is a universal quantifier, then the other is
	not a descendant of the former.
  \end{itemize}
\begin{proof}
By induction on $A$.
\begin{itemize}
  \item If $A$ is an atom, trivial.
  \item If $A = A_1 \cand A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are
	      adjacent in $\Gr(A)$ by definition. By Proposition \ref{prop1},
		    the one in $A_1$ (resp. $A_2$) is connected to the
		    vertex representing $A_1$ (resp. $A_2$) by a switching
		    path. Together with the two edges incident to $A_1 \cand
		    A_2$, we obtain a switching path connecting $u$ and $v$.
    \end{itemize}
  \item If $A = A_1 \cor A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are not
	      adjacent in $\Gr(A)$ by definition. It is clear that they
		    are not connected by a switching path.
    \end{itemize}
  \item If $A = \exists x \ A'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A'$: trivial by the induction
	      hypothesis.
      \item one of them is $\exists x$ and the other is in $A'$: trivial
	      by Proposition \ref{prop1}

@@

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = \Gr(A)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $A$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\cor$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\cor$: we can choose an
	    arbitrary switching for this $\cor$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching
		  path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes
		  through the left (resp. right) edge of $\cor$.

    \begin{center}	
    \begin{tikzpicture}
    \tikzstyle{every node}=[circle, minimum size=12pt, inner sep=2pt]
    \draw (0, 0) node (par) {$\cor$};
    \draw (1, 1.7) node (dots1) {$\vdots$};
    \draw (-1, 1.7) node (dots2) {$\vdots$};
    \draw (1, 3.4) node (vr) {$v_R$};
    \draw (-1, 3.4) node (vl) {$v_L$};
    \draw (1.5, -1) node (dots3) {$\vdots$};
    \draw (-1.5, -1) node (dots4) {$\vdots$};
    \draw (2.9, -1) node (ul) {$u_L$};
    \draw (-2.9, -1) node (ur) {$u_R$};
    \draw (par) -- (dots1) [color=blue];
    \draw (par) -- (dots2) [color=red];
    \draw (dots1) -- (vr) [color=blue]; 
    \draw (dots2) -- (vl) [color=red];
    \draw (dots4) -- (ur) [color=blue]; 
    \draw (dots3) -- (ul) [color=red];
    \draw (dots4) -- (par) [color=blue]; 
    \draw (dots3) -- (par) [color=red];
    \end{tikzpicture}

    {\bf Figure 7.} A schema showing that the two branches of the same $\cor$ cannot be used in the cycle at the same time. 
    \end{center}

@@

    It is clear that $u_L$ (resp. $u_R$) is not in the branches of the
	 $\cor$. Otherwise, there will be no switching path from $u_L$ to $v_L$

    By Proposition \ref{prop2}, we know that $u_L$ and $u_R$ are not
	universal quantifiers which are ancestors the $\cor$
	and that there exist one switching path from $u_L$ to $v_L$ and one
	from $u_R$ to $v_R$. In particular, there exist one switching path from
	$u_L$ to the $\cor$ and one from the $\cor$ to $v_R$, and by
	concatenating the two, we obtain a switching path from $u_L$ to $v_R$. 
	By Proposition \ref{prop2}, $u_L$ and $v_R$ are thus adjacent in $(W,
	E_G)$, which is impossible since $(W, E_W)$ is a matching.
  \end{itemize}
  Notice that the switching paths here are in the underlying formula tree. We
  have to verify that they are compatible with the choices of switching made for
  universal quantifiers. That is, if $uv \in E_W$, then for all the
  universal quantifiers $\forall x$ on the switching path (in the formula
	tree), we have chosen in the first part of the proof to keep the only
	edge from the child of $\forall x$ to itself. In fact, if there exists
	a universal quantifier $w \in W$ on the switching path $u \rightarrow
	v$, then one of $u$ and $v$ is not a descendant of $w$. Moreover, if
	$u$ (resp. $v$) is a universal quantifier, then $w$ is not in its scope.
	By Proposition \ref{prop2}, $\{wu, wv\} \cap E_W \neq \emptyset$, which
	is impossible since $(W, E_W)$ is a matching.
  We have thus constructed a switching graph containing this cycle.
\end{proof}
\end{proposition}

@@

\begin{definition}
Let $\theta$ be a unification structure on an $\FOMLL$ sequent $\Gamma$.
We define the \bfit{frame}  of $\theta$ by exhaustively applying the following subformula rewriting steps, to obtain a proof structure $\theta_m$ on an $\MLL$ sequent $\Gamma_m$:

\begin{enumerate}
  \item {\bf Encode dependencies as fresh links.} For each dependency $\exists x
	  \rightarrow \forall y$, with corresponding subformulas $\exists x A$
		and $\forall y B$, we add a fresh link as follows. Let $P$ be a
		fresh (nullary) predicate symbol. Replace $\exists x A$ with $P
		\cand \exists x A$ and $\forall y B$ with $\overline{P} \cor \forall y B$, and add an axiom link between $P$ and $\overline{P}$.
  \item {\bf Erase quantifiers.} After step 1, erase all the quantifiers. (We no longer need their leaps since they are encoded as links in step 1.
  \item {\bf Simplify atoms.} After step 2, replace every predicate $Pt_1 \cdots t_n$ with a nullary predicate symbol $P$.
\end{enumerate}
\end{definition}
